---
title: "Homework 3"
author: "Sophie Pardehpoosh"
date: "2026-02-26"
format:
  html:
    toc: true # use this to display a table of contents
execute:
  message: false # use this to make sure messages don't show up
  warning: false # use this to make sure warnings don't show up
---

## Github repository: https://github.com/spardehpoosh/ENVS-193DS_homework-03.git

# Set up

```{r}
#| label: reading in packages and data
#| message: false

library(tidyverse) # general use
library(janitor) # cleaning data frames
library(here) # file/folder organization
library(readxl) # reading .xlsx files

# reading in data
salinity <- read_csv(here("data", "salinity-pickleweed.csv")) # storing salinity pickleweed data as an object called salinity
reading_data <- read_csv(here("data", "personal_data.csv")) # storing personal data as an object called reading_data
```


# Problem 1. Slough soil salinity

## a. An apropriate test

To test the strength of the relationship between salinity and California pickleweed biomass, I would use Pearson's correlation or Spearman's Rank. Pearsons' correlation is the parametric version to measure the strength of a linear relationship. In contrast, Spearman rank correlation is the non-parametric version and tests for a monotonic relationship based on ranks between variables. Pearson's correlation requires an approximately normal distribution and continous variables while Spearman's rank does not have this requirement. 

## b. Create visualization 

```{r}
#| label: salinity scatterplot
#| message: false

# base layer: ggplot
ggplot(data = salinity, # using salinity data set for plot
       aes(x = salinity_mS_cm, # x-axis predictor - soil salinity
           y = pickleweed)) + # y-axis response - California pickleweed biomass
  geom_point(size = 4, # changing size of individual points
             stroke = 1, # making outline of points thickness 1
             fill = "magenta", # changing color of individual points
             shape = 21) + # changing shape of individual points
  labs(x = "Soil Salinity (mS/cm)", 
       y = "California pickleweed biomass (g)",
       title = "As soil salinity (mS/cm) increases, California pickleweed biomass (g) increases") + # adding title
  theme_minimal() # changing theme from ggplot default
```


## c. Check assumptions and run the test

### Check assumptions for Pearson's correlation

Both variables are assumed to be continuous by study design. 

Based on the scatterplot in part 1b, there seems to be a linear relationship between soil salinity (mS/cm) and California pickleweed biomass (g). 

#### Model fitting: 
```{r}
#| label: model fitting

salinity_model <- lm(
  pickleweed ~ salinity_mS_cm, # formula: change in California pickleweed biomass (g) as a function of soil salinity (mS/cm)
  data = salinity) # salinity dataset

```

#### Diagnostics 
```{r}
#| label: diagnostics

par(mfrow = c(2, 2)) # display residual plots in 2x2 grid
plot(salinity_model) # plot residuals

```
We evaluated model assumptions visually. For homoscedasticity of residuals we used Residuals vs. Fitted and Scale-Location plots. For normality of residuals we used QQ plots. Based on our diagnostics, there were no major deviations from homoscedasticity or normality of residuals. There were also no outliers that might influence model estimates or predictions.

#### Normality
```{r}
#| label: qqplot for pickleweed biomass
#base layer: ggplot
ggplot(data = salinity, # starting with the clean lobsters data frame
       aes(sample = salinity_mS_cm, # y-axis for gg plot
           )) + 
  geom_qq_line(color = "maroon4") + # reference line in maroon
  geom_qq() + # points on qqplot
  theme(legend.position = "none") # removing the legend
```

```{r}
#| label: qqplot for pickleweed
#base layer: ggplot
ggplot(data = salinity, # starting with the clean lobsters data frame
       aes(sample = pickleweed, # y-axis for gg plot
           )) + 
  geom_qq_line(color = "maroon4") + # reference line in maroon
  geom_qq() + # points on qqplot
  theme(legend.position = "none") # removing the legend
```

Looking at the qqplot, it seems as soil salinity and pickle weed biomass are normally distributed because the points on both QQplot follow the straight line, especially in the middle of the line

### Run the test

```{r}
#| label: running the test

cor.test(salinity$salinity_mS_cm, salinity$pickleweed, method = "pearson") # pearsons test for strength of relationship between salinity and pickleweed biomass

```

## d. Results communication 



## e. Test implications

## f. Double check your work


# Problem 2. Personal data 

## a. Updating your visualizations 

## b. Captions


# Problem 3. Affective visualization 

## a. Description of visualization


## b. Sketch of visualization


## c. Draft of visualization 

## d. Artist statement


## e. Prep materials to share in class


# Problem 4. Statistical critique

## a. Revisit and summarize


## b. Visual clarity


## c. Aesthetic clarity


## d. Recommendations





